---
title: "Assignment 1"
author: "Manuel"
date: "8 Februar 2017"
output: html_document
fig_caption: true
---

============

## Loading and preprocessing the data

Load packages and read data.

```{r}

library(tidyverse)
activity <- read_csv("H:/Kurse/DataScience/ReproducibleResearch/Assignment1/activity.csv", col_types = cols(steps = col_number()))

```

## What is mean total number of steps taken per day?

Group by date and calculate sum of steps per day

```{r}
sum_steps <- activity %>% group_by(date) %>% summarise (sum = sum (steps))
```

Make a histogram of the sum of steps per day

```{r}
hist(sum_steps$sum)
```

Calculate the mean and median steps per day

```{r}
sum_steps %>% group_by(date) %>% summarise (mean = mean(sum), median = median (sum, na.rm = TRUE))
```

## What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all days (y-axis)

# Calculate the mean of steps per interval across all days

```{r}
intervals <- activity %>% group_by(interval) %>% summarise (mean=mean(steps, na.rm = TRUE))

ggplot(intervals, aes(x =  interval, y = mean)) +
  geom_line() +
  theme_bw()

```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
intervals %>% arrange(desc(mean))
```

## Imputing missing values

Calculate NAs

```{r}
sum(is.na(activity$steps))
```

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean for that 5-minute interval. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity_imputed <- activity %>% group_by(interval) %>% mutate(steps = replace(steps, is.na(steps), mean(steps, na.rm=TRUE)))
sum(is.na(activity_imputed$steps))
```

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
sum_steps2 <- activity_imputed %>% group_by(date) %>% summarise (sum = sum (steps))
hist(sum_steps2$sum)
```

Answer: Imputing changes results significantly.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activity_imputed <- activity_imputed %>% mutate(date, Weekday = weekdays(date))
activity_imputed <- activity_imputed %>% mutate(Weekday = ifelse(Weekday == "Sonntag" | Weekday == "Samstag", "Weekend", "Workday"))
```

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
interval2 <- activity_imputed %>% group_by(Weekday, interval) %>% summarise (mean = mean (steps))

ggplot(interval2, aes ( x = interval, y = mean, color = Weekday)) +
         geom_line(size = 0.8) +
         facet_wrap(~Weekday, ncol = 1, nrow=2) +
         theme_bw()
```